---
title: "Lecture 10: Law of Large Numbers"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
set.seed(237)
```

## R prepartion

### `ggplot` and colours

`ggplot2` allows users to easily customize the visual properties of the plots
including the colours. You can assign colours to `fill` and `color` arguments
inside the functions adding the geometry objects - e.g., `geom_point()`, `geom_line()`, etc.
and annotations.

We have practiced using color in Weekly Activity 3.


```{r}
library(ggplot2)
set.seed(123)
N <- 101
theta <- 1/4
sims <- data.frame(
  index = seq_len(N),
  value = sort(rgeom(N, theta) + 1)
)
# Here, I am using a data frame inside ggplot()
sims_base <- ggplot(sims, aes(x = index, y = value)) +
  theme_classic() +
  labs(x = "Cumulative count", y = "Simulated value") +
  scale_x_continuous(breaks = c(1, 26, 51, 76, 101)) +
  scale_y_continuous(
    breaks = c(range(sims$value), median(sims$value), 1 / theta)
  )
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta), 
               color = "darkgrey", linewidth = 1) +
  geom_point(size = 2)  +
  geom_hline(yintercept = 1 / theta, linetype = "solid", 
             linewidth = 2, color = "navy") +
  geom_hline(yintercept = mean(sims$value), linetype = "dashed", 
             linewidth = 1, color = "orange") +
  annotate("segment", x = 51, xend = 51, y = 1, yend = sims$value[51],
           linetype = "dotted", color = "red", linewidth = 1) +
  annotate("segment", x = 0, xend = 51, y = sims$value[51], yend = sims$value[51],
           linetype = "dotted", color = "red", linewidth = 1) +
  annotate("text", x = c(1, 30, 52), y = c(4.5, 4.5, 3), 
           label = c("Simulated mean", "E[G]", "Simulated median"),
           vjust = c(0, 0, 1.5), hjust = 0, color = c("navy", "orange", "red"))
```

`ggplot2` also allows you to map data values to colors of geometry objects. You
can define the mappings inside `aes()` calls. The example below colors the line
segments and points based on whether the `value` is less than `1 / theta` or not.

```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2)
```

`ggplot2` uses a set of default colours and creates a legend. To control the
specific colours mapped to data values and the legend, you can use the
family of `scale_colour_*()` functions. For example, we can set the name to be
blank and use descriptive labels for the legend keys.


```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2) +
  scale_colour_discrete(name = "", labels = c("Not below mean", "Below mean"))
```

> We used `scale_colour_discrete()` because the mapped data is of type logical and
`ggplot2` considers data types logical and characters as "discrete". Numerics are considered
as "continuous". They are based on data types coded in R and may be different from 
the context. e.g., a count may be coded as a "continuous" data type. We will look at
the case of "discrete" data types.

There are a few options to control the specific colours mapped to data.

+  `scale_colour_manual()` allows you to map a specific colour to each value
manually using `values` argument. Common colour names such as "black", "orange",
etc. are accepted or you can use [Hex color codes](https://www.color-hex.com).

```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2) +
  scale_colour_manual(
    # map colours to values
    values = c(`TRUE` = "orange", `FALSE` = "grey"),
    name = "", labels = c("Not below mean", "Below mean"),
  )
```


+  `scale_colour_brewer()` allows you to use preset colour schemes from
[ColorBrewer](https://colorbrewer2.org). You can select a colour scheme by type 
(`type = "div"` for diverging colour scheme) or by a specific name (`palette = "Pastel1"`).

```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2) +
  scale_colour_brewer(
    type = "div", # select a diverging colour scheme
    name = "", labels = c("Not below mean", "Below mean"),
  )
```

```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2) +
  scale_colour_brewer(
    palette = "Pastel1", # select a diverging colour scheme
    name = "", labels = c("Not below mean", "Below mean"),
  )
```

+  `scale_viridis_d()` provides preset colour schemes called [viridis](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html).
They are designed to be easy on the eay and colour-blind friendly.

```{r}
sims_base +
  geom_segment(aes(xend = index, yend = 1 / theta,
                   color = value < (1 / theta))) +
  geom_point(aes(color = value < 1 / theta), size = 2) +
  scale_colour_viridis_d(
    name = "", labels = c("Not below mean", "Below mean"),
  )
```


```{r color-ex-setup}
set.seed(123)
N <- 101
theta <- 1/4
sims <- data.frame(
  index = seq_len(N),
  value = sort(rgeom(N, theta) + 1)
)
# Here, I am using a data frame inside ggplot()
sims_base <- ggplot(sims, aes(x = index, y = value)) +
  theme_classic() +
  labs(x = "Cumulative count", y = "Simulated value") +
  scale_x_continuous(breaks = c(1, 26, 51, 76, 101)) +
  scale_y_continuous(
    breaks = c(range(sims$value), median(sims$value), 1 / theta)
  )
```


In the code chunk below, the `sims` and `sims_base` objects are available. A points at 
coordinates `(index, value)` with `shape = 21`. The shape allows your to
control `fill` for the filling colour and `colour` for the edge colour separately.
Set `fill` based on `value < sims$value[51]` using ColorBrewer palette _Dark2_.


```{r color-ex, exercise=TRUE, exercise.lines=10}
sims_base
```

```{r color-ex-solution}
sims_base +
  geom_point(aes(fill = value < sims$value[51]), shape = 21) +
  scale_fill_brewer(palette = "Dark2", name = "", 
                      labels = c("Not below sample median", "Below sample median"))
```


## The law of large numbers

For each of the distributions below, complete the following steps to investigate
the effect of the law of large numbers on their sample means.

1. Simulate $1\ 000$ random samples, $X_i$ for $i=1,2,\ldots,1\ 000$.
2. Compute sample means for the first $n$ samples for $n=1,2,\ldots,1\ 000$. That is,
$$\overline{X}_n=\frac{1}{n}\sum_{i=1}^n X_i$$
3. Plot the sample means in order as points using different colours
for those within $1/2$ of the population standard deviation from the mean. Use colours of your choice.

### Normal Distribution

$$N(0,1)$$

```{r normal-lln, exercise=TRUE, exercise.lines=15}
N <- 1000
x <- rnorm(N)
sims <- numeric(N)
for (n in seq_len(N)) {
  sims[n] <- sum(x[1:n]) / n
}
```

```{r normal-lln-solution}
N <- 1000
x <- rnorm(N)
sims <- numeric(N)
for (n in seq_len(N)) {
  sims[n] <- sum(x[1:n]) / n
}
ggplot() +
  theme_classic() +
  geom_point(aes(x = seq_len(N), y = sims, colour = abs(sims) < 1 / 2)) +
  scale_colour_manual(values = c(`FALSE` = "navy", `TRUE` = "pink"),
                      labels = c(`FALSE` = "NO", `TRUE` = "YES"),
                      name = "Within half SD from mean")
```


### Exponential distribution

$$\text{Exp}(1)$$


```{r exp-lln, exercise=TRUE, exercise.lines=15}
N <- 1000

```

```{r exp-lln-solution}
N <- 1000
x <- rexp(N)
sims <- numeric(N)
for (n in seq_len(N)) {
  sims[n] <- sum(x[1:n]) / n
}
ggplot() +
  theme_classic() +
  geom_point(aes(x = seq_len(N), y = sims, colour = abs(sims - 1) < 1 / 2)) +
  scale_colour_manual(values = c(`FALSE` = "navy", `TRUE` = "pink"),
                      labels = c(`FALSE` = "NO", `TRUE` = "YES"),
                      name = "Within half SD from mean")
```


### Poisson distribution

$$\text{Pois}(1)$$


```{r pois-lln, exercise=TRUE, exercise.lines=15}
N <- 1000

```

```{r pois-lln-solution}
N <- 1000
x <- rpois(N, 1)
sims <- numeric(N)
for (n in seq_len(N)) {
  sims[n] <- sum(x[1:n]) / n
}
ggplot() +
  theme_classic() +
  geom_point(aes(x = seq_len(N), y = sims, colour = abs(sims - 1) < 1 / 2)) +
  scale_colour_manual(values = c(`FALSE` = "navy", `TRUE` = "pink"),
                      labels = c(`FALSE` = "NO", `TRUE` = "YES"),
                      name = "Within half SD from mean")
```

### Binomial Distribution

$$\text{Binom}\left(10, 1/10\right)$$

```{r binom-lln, exercise=TRUE, exercise.lines=15}
N <- 1000

```

```{r binom-lln-solution}
N <- 1000
x <- rbinom(N, 10, 1 / 10)
sims <- numeric(N)
for (n in seq_len(N)) {
  sims[n] <- sum(x[1:n]) / n
}
ggplot() +
  theme_classic() +
  geom_point(aes(x = seq_len(N), y = sims, 
                 colour = abs(sims - 1) < sqrt(9 / 10) / 2)) +
  scale_colour_manual(values = c(`FALSE` = "navy", `TRUE` = "pink"),
                      labels = c(`FALSE` = "NO", `TRUE` = "YES"),
                      name = "Within half SD from mean")
```
