---
title: "Lecture 10: Law of Large Numbers"
output: 
  learnr::tutorial:
    css: css/styles.css
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
library(tidyverse)
library(ggplot2)
library(kableExtra)
```


## Law of Large Numbers

### Normal Distribution

Simulate 1,000 samples of $X\sim N(0,1)$.

#### Population vs. Sample Distribution

Plot histograms for the first 10, 100, and 1000 samples. Fix the x-axis range to -5 to 5.

```{r hist1, exercise=TRUE, exercise.lines=5}
x <- rnorm(1000)
```

Compare them to the histograms from lecture slides (Example: Average Samples).

#### Law of Large Numbers

b) Compute

$$\overline{X}_n=\frac{1}{n}\sum_{i=1}^n X_i$$

for $n=1,2,\dots,1000$ and plot them as points along y-axis with $n$ on x-axis.


```{r trace1, exercise=TRUE, exercise.lines=5}
x <- rnorm(1000)
```

### Other Distributions

Repeat the exercise with $X\sim\text{Pois}(3)$.

```{r trace2, exercise=TRUE, exercise.lines=25}
```

Repeat the exercise with $Y\sim\text{Exp}(0.5)$


```{r trace3, exercise=TRUE, exercise.lines=25}
```

## Sampling distributions

For each random variable below, generate 10,000 simulations of $\overline{X}_{10}$, $\overline{X}_{20}$, $\overline{X}_{30}$, and $\overline{X}_{50}$, where $\overline{X}$ is the average of $n$ independent samples of $X$. Compare their distributions by plotting `density()` on a single plot.

### Normal Distrbution

$$X\sim N(5, 5^2)$$

```{r echo=FALSE, out.width="50%", out.height="50%"}
ggplot() +
  theme_classic() +
  geom_function(fun = dnorm, args = list(mean = 5, sd = 5)) +
  scale_x_continuous(limits = c(-20, 30)) +
  labs(y = "f", title = "Probability density funtion of N(5,25)")
```


```{r norm, exercise=TRUE, exercise.lines=20}
n <- 10000
xbar10 <- replicate(n, mean(rnorm(10, 5, 5)))
```

### Exponential Distribution

$$X\sim \text{Exp}\left(1/5\right)$$

```{r echo=FALSE, out.width="50%", out.height="50%"}
ggplot() +
  theme_classic() +
  geom_function(fun = dexp, args = list(rate = 0.2)) +
  scale_x_continuous(limits = c(0, 30)) +
  labs(y = "f", title = "Probability density funtion of Exp(1/5)")
```


```{r exp, exercise=TRUE, exercise.lines=20}
n <- 10000
xbar10 <- replicate(n, mean(rexp(10, 1/5)))
```

### Binomial Distribution

$$X\sim \text{Binom}\left(10, 1/10\right)$$

```{r echo=FALSE, out.width="50%", out.height="50%"}
r <- 0:10
db <- dbinom(r, 10, 1/10)
ggplot() +
  theme_classic() +
  geom_segment(aes(x = r, xend = r, y = db, yend = 0), linetype = "dotted") +
  geom_point(aes(x = r, y = db)) +
  labs(y = "p", title = "Probability mass funtion of Binom(10, 1/10)")
```

```{r binom, exercise=TRUE, exercise.lines=20}
n <- 10000
xbar10 <- replicate(n, mean(rbinom(10, 10, 1/10)))
```


### Poisson Distribution

$$X\sim \text{Pois}\left(10\right)$$

```{r echo=FALSE, out.width="50%", out.height="50%"}
r <- 0:50
db <- dpois(r, 10)
ggplot() +
  theme_classic() +
  geom_segment(aes(x = r, xend = r, y = db, yend = 0), linetype = "dotted") +
  geom_point(aes(x = r, y = db)) +
  labs(y = "p", title = "Probability mass funtion of Pois(10)")
```

```{r pois, exercise=TRUE, exercise.lines=20}
n <- 10000
xbar10 <- replicate(n, mean(rpois(10, 10)))
```
